---
title: "tuner-app"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{tuner-app}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction
  
When students submit R scripts as homework, they will probably be required 
to read in some data, perform some analysis and create some plots.

Many times, students use relative paths incorrectly, or use absolute paths. This
prevents their code from running on our machines. This app was written in the
spirit of getting students to write code that **others** can run.

To try out the app, you can work with the two solution templates installed 
with the package. It will use whatever your current working directory as the 
working directory, so do make sure you have write-access to it.

```{r tuner_demo, eval=FALSE}
library(autoharp)
run_tuner("demo", system.file("examples", "soln_templates", normalizePath("."))
```


